Process Genotypes For Ancestry Analysis

### 2019-09-27 code/Ancestry_PCA.Rmd
Break Out Python

```{r libs, include=FALSE}
knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  results='hide')
library(reticulate)
```
##Before running 
Run the following in docker exec -it <CONTAINER> /bin/bash 
chmod 777 /root
mv TWAS/ /home/<USR>/TWAS/
chown -hR <GID>:<USR> /home/<USR>/TWAS/

@@

#Alternativly you can setup a credentials file as such:
touch ~/.synapseConfig
echo "[authentication]" >> ~/.synapseConfig
echo "username = <USR>" >> ~/.synapseConfig
echo "password = <PASWD>" >> ~/.synapseConfig
```
## Pull 1000Genomes down and process
* Pull out reference populations (CEU, YRI, CHB, JPT)
```{bash pull_1000Genomes, echo=T, results='hide', cache = FALSE, echo=FALSE}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
#PATH=$PATH:TWAS/bin/
CORES=`expr $(nproc) - 2`

@@

#Convert to PLINK
for i in {22..1}
do
  plink --vcf VCF/NIA_JG_1898_samples_GRM_WGS_b37_JointAnalysis01_2017-12-08_$i\.recalibrated_variants.MSSM.vcf.gz --threads $CORES --make-bed --silent --allow-no-sex --snps-only --out GWAS_Plink/MSSB_Chr_$i
  rm VCF/NIA_JG_1898_samples_GRM_WGS_b37_JointAnalysis01_2017-12-08_$i\.recalibrated_variants.MSSM.vcf.*
done
rm -r VCF/

@@

plink --threads $CORES --bfile GWAS_Plink/MSSB_Chr_1 --silent --merge-list MergedGWAS.txt --extract MSBB_LD_SNPS.txt --make-bed --allow-no-sex --out GWASTotal
#rm -r GWAS_Plink/
rm MergedGWAS.txt

@@

### ReName the MSBB SNPS so the can be combined
```{python ReNamer, echo=T, results='hide'}
import sys
import os
import re

InFile = 'TotalLDRef.bim'
InFileb = 'GWASTotal.bim' #LDREF_FiltSNPs.bim

RS_IDs = {}
RS_Pos = {}

F1 = open(InFile)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	RS_IDs[ ''.join([lst[0], '_', lst[3]]) ] = lst[1]
	RS_Pos[ ''.join([lst[0], '_', lst[3]]) ] = lst[2]
F1.close()

#Open Output
OUT = open( ''.join([ InFileb.replace(".bim", ""), 'Renamed.bim']), 'w' )

F1 = open(InFileb)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	Name = RS_IDs[ ''.join([ lst[0], "_", lst[3] ])]
	Post = RS_Pos[ ''.join([ lst[0], "_", lst[3] ])]

	Entry = '\t'.join([ lst[0], Name, Post, lst[3], lst[4], lst[5] ])
	print >>OUT, Entry

F1.close()
OUT.close()
```

@@

##This Number should be zero:
Repod=$(awk '{print $2}' GWASTotalRenamed.bim | grep "\." - | wc -l)
if [ "$Repod" -eq "0" ]; then
   mv GWASTotalRenamed.bim GWASTotal.bim;
fi
if [ "$Repod" -ne "0" ]; then
   echo $Repod" lines could've failed Translation"
fi
plink --bfile GWASTotal --silent --threads $CORES --make-bed --out MSBB_Preliminary_Genos
rm GWASTotal.*
```

@@

### ReName the Mayo SNPS so the can be combined
```{python ReNamerMayo, echo=T, results='hide'}
import sys
import os
import re

InFile = 'TotalLDRef.bim'
InFileb = 'Mayo_LDREF_Filt.bim'

RS_IDs = {}
RS_Pos = {}

F1 = open(InFile)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	RS_IDs[ ''.join([lst[0], '_', lst[3]]) ] = lst[1]
	RS_Pos[ ''.join([lst[0], '_', lst[3]]) ] = lst[2]
F1.close()

#Open Output
OUT = open( ''.join([ InFileb.replace(".bim", ""), 'Renamed.bim']), 'w' )

F1 = open(InFileb)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	Name = RS_IDs[ ''.join([ lst[0], "_", lst[3] ])]
	Post = RS_Pos[ ''.join([ lst[0], "_", lst[3] ])]

	Entry = '\t'.join([ lst[0], Name, Post, lst[3], lst[4], lst[5] ])
	print >>OUT, Entry

F1.close()
OUT.close()
```

@@

### ReName the Mayo SNPS so the can be combined
```{python ReNamerRosMap, echo=T, results='hide'}
import sys
import os
import re

InFile = 'TotalLDRef.bim'
InFileb = 'ROSMAP_temp.bim'

RS_Chr = {}
RS_Pos = {}
RS_BP ={}

F1 = open(InFile)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')

	RS_Chr[ ''.join([lst[1], '_', lst[4], '_', lst[5]]) ] = lst[0]
	RS_Pos[ ''.join([lst[1], '_', lst[4], '_', lst[5]]) ] = lst[2]
	RS_BP[ ''.join([lst[1], '_', lst[4], '_', lst[5]]) ] = lst[3]
F1.close()

#Open Output
OUT = open( ''.join([ InFileb.replace(".bim", ""), 'Renamed.bim']), 'w' )
OUT2 = open( 'FailedSNPS.txt', 'w' )

F1 = open(InFileb)
for line in F1:
	LINE = line.rstrip('\r\n')
	lst = LINE.split('\t')
	LOC = ''.join([ lst[1], '_', lst[4], '_', lst[5] ])
	chrm = RS_Chr.get( LOC, None )
	pos = RS_Pos.get( LOC, None )
	bp = RS_BP.get( LOC, None )

	if( chrm == None ):
	  print >>OUT, LINE
	  print >>OUT2, lst[1]
	else:
	  Entry = '\t'.join([ chrm, lst[1], pos, bp, lst[4], lst[5] ])
	  print >>OUT, Entry

F1.close()
OUT.close()
OUT2.close()
```

@@

rm peds.txt
rm Names.txt
rm PEDS.txt
AllSnps4GemClust.*
```

##Plot PCA -- Needs Work...
```{r PlotPCA, cache = FALSE, echo=FALSE}
library(ggplot2)
library(plotrix)
library(scater)
library(data.table)
source('GemTools.R')

@@

mydf$task <- factor(mydf$task, levels = c("up", "down", "left", "right", "front", "back"))
P <- ggplot(pcs, aes( PC1, PC2) ) + geom_point( aes(colour = Ancestry, shape = SampleType), cex=.8, alpha = 4/10) 
P <- P + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
P <- P + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" ))  
P <- P + ggtitle("Ancestry Cluster Analysis CEU vs NonCEU") + theme(plot.title = element_text(hjust = 0.5))
P

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn20798011';
activityName = 'Ancestry Analysis';
activityDescription = 'Processing and Analysis of WGS and Genotype data along with ancestry analysis with 1000 genome\'s CEU, CHB, YRI, and JPT populations';
thisFileName <- 'Ancestry_PCA.Rmd'
# Github link
thisRepo <- getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

```{r Store, cache = FALSE, echo=FALSE, results='hide'}
parentId = 'syn20798011';
activityName = 'Ancestry Analysis';
activityDescription = 'Processing and Analysis of WGS and Genotype data along with ancestry analysis with 1000 genome\'s CEU, CHB, YRI, and JPT populations';

@@

#STORE
#CEU Indvs
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryClustered_INDVs.tsv', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryClustered_INDVs.tsv', 
                  name = 'Samples Clustered to CEU Ancestry', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'CEU_INDV'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#PCA EginVecs
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'Total_PCA.eigenvec', sep = '\t', quote=F)
ENRICH_OBJ = File('Total_PCA.eigenvec', 
                  name = 'PCA Eigen Vectors', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'EigenVecs'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#PCA EginVecs
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'Total_PCA.eigenval', sep = '\t', quote=F)
ENRICH_OBJ = File('Total_PCA.eigenval', 
                  name = 'PCA Eigen Values', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'EigenValues'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

@@

#Geno Bed
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.bed', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.bed', 
                  name = 'Genotype Bed File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno Bim
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.bim', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.bim', 
                  name = 'Genotype Bim File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno Fam
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.fam', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.fam', 
                  name = 'Genotype Fam File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno nosex
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.nosex', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.nosex', 
                  name = 'Genotype nosex File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno nosex
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.log', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.log', 
                  name = 'Genotype log File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)
```
### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
library(synapser)
library(knit2synapse) 
#install.packages("reticulate")
library(reticulate)

# get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "Ancestry_PCA.Rmd",
                                          parentId ="syn20798011",
                                          folderName = 'Ancestry_Analysis')

>>>>

```{r libs, include=FALSE}
knitr::opts_chunk$set(
  #python.reticulate=FALSE,
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  results='hide')
```
##Before running 
Run the following in docker exec -it <CONTAINER> /bin/bash 
chmod 777 /root
mv TWAS/ /home/<USR>/TWAS/
chown -hR <GID>:<USR> /home/<USR>/TWAS/
maybe# ln -s /usr/bin/python /usr/local/bin/python

@@

#Alternativly you can setup a credentials file as such:
touch ~/.synapseConfig
echo "[authentication]" >> ~/.synapseConfig
echo "username = <USR>" >> ~/.synapseConfig
echo "password = <PASWD>" >> ~/.synapseConfig
```

## Pull 1000Genomes down and process
* Pull out reference populations (CEU, YRI, CHB, JPT)
```{bash pull_1000Genomes, echo=T, results='hide', cache = FALSE}
source /root/.bashrc
PATH=$PATH:/home/jgockley/TWAS/TWAS/bin/
#PATH=$PATH:TWAS/bin/
CORES=`expr $(nproc) - 2`

@@

#Convert to PLINK
for i in {22..1}
do
  plink --vcf VCF/NIA_JG_1898_samples_GRM_WGS_b37_JointAnalysis01_2017-12-08_$i\.recalibrated_variants.MSSM.vcf.gz --threads $CORES --make-bed --silent --allow-no-sex --snps-only --out GWAS_Plink/MSSB_Chr_$i
  #rm VCF/NIA_JG_1898_samples_GRM_WGS_b37_JointAnalysis01_2017-12-08_$i\.recalibrated_variants.MSSM.vcf.*
done
#rm -r VCF/

@@

plink --threads $CORES --bfile GWAS_Plink/MSSB_Chr_1 --silent --merge-list MergedGWAS.txt --extract MSBB_LD_SNPS.txt --make-bed --allow-no-sex --out GWASTotal
#rm -r GWAS_Plink/
#rm MergedGWAS.txt

@@

### ReName the MSBB SNPS so the can be combined
```{r ReNamer_ONE, echo=F, results='hide'}
system("python Python/First_Rename.py")
```

@@

##This Number should be zero:
Repod=$(awk '{print $2}' GWASTotalRenamed.bim | grep "\." - | wc -l)
if [ "$Repod" -eq "0" ]; then
   cp GWASTotalRenamed.bim GWASTotal.bim;
fi
if [ "$Repod" -ne "0" ]; then
   echo $Repod" lines could've failed Translation"
fi
plink --bfile GWASTotal --silent --threads $CORES --make-bed --out MSBB_Preliminary_Genos
#rm GWASTotal.*
```

@@

### ReName the Mayo SNPS so the can be combined
```{r ReNamer_Two_Mayo, echo=T, results='hide'}
system("python Python/Second_Rename.py")
```

@@

### ReName the Mayo SNPS so the can be combined
```{r ReNamer_Three_Rosmap, echo=T, results='hide'}
system("python Python/Third_Rename.py")
```

@@

rm peds.txt
AllSnps4GemClust.*
```

##Plot PCA
```{r PlotPCA, cache = FALSE, echo=FALSE}
library(ggplot2)
library(plotrix)
library(scater)
library(data.table)
source('GemTools.R')

@@

#mydf$task <- factor(mydf$task, levels = c("up", "down", "left", "right", "front", "back"))
P <- ggplot(pcs, aes( PC1, PC2) ) + geom_point( aes(colour = Ancestry, shape = SampleType), cex=.8, alpha = 4/10) 
P <- P + xlab(paste0( "PC1 ", signif( PC1_Var_exp*100, 3), "%" )) 
P <- P + ylab(paste0( "PC2 ", signif( PC2_Var_exp*100, 3), "%" ))  
P <- P + ggtitle("Ancestry Cluster Analysis CEU vs NonCEU") + theme(plot.title = element_text(hjust = 0.5))
P

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
library(githubr)
parentId = 'syn18936948';
activityName = 'Ancestry Analysis';
activityDescription = 'Processing and Analysis of WGS and Genotype data along with ancestry analysis with 1000 genome\'s CEU, CHB, YRI, and JPT populations';
thisFileName <- 'Ancestry_PCA.Rmd'
# Github link
thisRepo <- getRepo(repository = "jgockley62/TWAS", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/',thisFileName))
```

```{r Store, cache = FALSE, echo=FALSE, results='hide', eval=FALSE}
parentId = 'syn18936948';
activityName = 'Ancestry Analysis';
activityDescription = 'Processing and Analysis of WGS and Genotype data along with ancestry analysis with 1000 genome\'s CEU, CHB, YRI, and JPT populations';

@@

#STORE
#CEU Indvs
#write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryClustered_INDVs.tsv', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryClustered_INDVs.tsv', 
                  name = 'Samples Clustered to CEU Ancestry', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'CEU_INDV'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#PCA EginVecs
#write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'Total_PCA.eigenvec', sep = '\t', quote=F)
ENRICH_OBJ = File('Total_PCA.eigenvec', 
                  name = 'PCA Eigen Vectors', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'EigenVecs'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#PCA EginVecs
#write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'Total_PCA.eigenval', sep = '\t', quote=F)
ENRICH_OBJ = File('Total_PCA.eigenval', 
                  name = 'PCA Eigen Values', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'EigenValues'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

@@

#Geno Bed
#write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.bed', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.bed', 
                  name = 'Genotype Bed File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno Bim
#write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.bim', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.bim', 
                  name = 'Genotype Bim File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno Fam
#write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.fam', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.fam', 
                  name = 'Genotype Fam File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno nosex
#write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.nosex', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.nosex', 
                  name = 'Genotype nosex File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Geno log
#write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CEU_AncestryMatched_Genotypes.log', sep = '\t', quote=F)
ENRICH_OBJ = File('CEU_AncestryMatched_Genotypes.log', 
                  name = 'Genotype log File', 
                  parentId = CODE$properties$id)
ENRICH_OBJ = synStore(ENRICH_OBJ, used = Syns_Used, activityName = activityName, 
                      executed = thisFile, activityDescription = activityDescription)
all.annotations$dataSubType = 'GWAS'
synSetAnnotations(ENRICH_OBJ, annotations = all.annotations)
```
### R Source Code
[Github](`r thisFile`)

```{r knitmd, eval=FALSE, cache=FALSE, include=FALSE}
library(synapser)
library(knit2synapse)

synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "Ancestry_PCA.Rmd",
                                          parentId ="syn18936948",
                                          folderName = 'Ancestry_Analysis')
